---
title: "Alignment of slices from a single patient"
author: "Victor Gaya"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alignment of slices from a single patient}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Loading required packages
```{r}
#| label: packages
#remove.packages("STIMA")
#remotes::install_github("vagm110901/STIMA")
library(STIMA)
```


## Defining input directory
```{r, eval=FALSE}
#| label: folder
# Define the folder where raw data is located
file_url <- "https://github.com/vagm110901/STIMA-exampleData/raw/refs/heads/master/inst/extdata/exampleData/Paciente19_merge.rds"
dest_dir <- tempdir()
dest <- file.path(dest_dir, "Paciente19_merge.rds")
utils::download.file(file_url, destfile = dest, mode = "wb")

inputDir <- dest_dir

paciente_merge <- readRDS(dest)
```

## Perform the alignment
```{r, eval=FALSE}
#| label: STIMA_1
listAlignedGTEM <- STIMA::STIMA(paciente_merge, mode = "GTEM", scale = FALSE)
listAlignedprocrustes <- STIMA::STIMA(paciente_merge, mode = "procrustes", scale = FALSE)
listAlignedRVSSimageJ <- STIMA::STIMA(paciente_merge, mode = "RVSSimageJ", scale = FALSE)
```

```{r, eval=FALSE}
#| label: data
listAlignedGTEM <- readRDS(paste0(inputDir,"/results/objectAligned_merge_GTEM.rds"))
```


```{r, eval=FALSE}
#| label: STIMA_2
paciente_merge_aligned <- listAlignedGTEM$alignedObj
listCoordinates <- listAlignedGTEM$listCoord
listCoordinatesNew <- listAlignedGTEM$listCoordNew
```

## Evaluation of the alignment
```{r, eval=FALSE}
#| label: Evalaution
# Do it for each method
EvaluationGTEM <- STIMA::calculateEvaluation(paciente_merge_aligned, mode = "GTEM",
                                             listCoordinatesNew, listCoordinates,
                                             patientType = "unique")
```

## Deconvolution
Creation of the deconvolution objects with three parts: 
  the reference object
  the problem object before the alignment 
  the aligned problem object
```{r, eval=FALSE}
#| label: Deconvolution_1
# Do it for each method
STIMA::createDeconvolutionLists(paciente_merge_aligned, mode = "GTEM")
```

Perform the deconvolution.
```{r, eval=FALSE}
#| label: Deconvolution_2
# Do it for each method and image
file_url <- "https://github.com/vagm110901/STIMA-exampleData/raw/refs/heads/master/inst/extdata/exampleData/snRNA_QC_Yadav.rds"
dest_dir <- tempdir()
dest_ref <- file.path(dest_dir, "snRNA_QC_Yadav.rds")
utils::download.file(file_url, destfile = dest_ref, mode = "wb")

reference <- readRDS(dest_ref)

paciente_merge_aligned_list_im2 <- readRDS("./results/objectAligned_merge_GTEM_list_im2.rds")
STIMA::deconvolutionRCTD(paciente_merge_aligned_list_im2, reference, 
                         im = "2", mode = "GTEM")

```

Merge of the deconvolution resutls for each image in an individual list object.
```{r, eval=FALSE}
#| label: Deconvolution_3
listDeconv <- STIMA::deconvolutionRCTD_mergeFiles(modes = c("GTEM","procrustes"), ims = c("2","3","4"))
```

Calculate the RV coefficients
```{r, eval=FALSE}
setwd("/Users/vagm_110901/Documents/Universitat/ConesaLab/Lab")
RVcoeff <- STIMA::calculateRVcoeff(modes = c("GTEM","procrustes"), ims = c("2","3","4"))
```
